### 2020-08-22 article/Article_meta_en.Rmd
Minor fixed in references and article texts

---
title: "An assessment of openness, availability and reliability of official information on sulphur dioxide emissions from the largest anthropogenic point sources in Russia using remote sensing data"
author: "P. O. Syomin"
affiliation: "Limited Liability Company Habr, 2/1 Spartakovskiy Lane, Section 3, Moscow, 105082, Russian Federation"
abstract: "Everyone has a right to information about the state of the environment. The purpose of the study is to assess the openness, availability and reliability of official information on sulphur dioxide emissions from 31 of the largest point anthropogenic sources located in Russia. For the assessment, a comparison was made of the amounts of emissions according to official reporting and remote sensing data. The set of reported data was formed using information from websites of Federal State Statistic Service and of Federal Service for Supervision of Natural Resources, as well as those contained in state reports on the ecological situation and on environmental protection or in public non-financial statements of large companies. Remote sensing dataset was downloaded from NASA Global Sulfur Dioxide Monitoring project website. Data comparison was carried out using statistical analysis methods. It is found that remote sensing data can be considered a standard of information openness and availability, while reported data is incomplete and require significant efforts to collect. For the emissions up to 50 000 tons per year, the differences between the reported and remote sensing data can be explained solely by the uncertainty of satellite measures. With a larger amount, however, the remote sensing values are in general larger than reported ones, which may indicate imperfect systems of emission control from stationary sources, the presence of many unaccounted sources, or deliberate distortion of environmental information. Emission catalogues from satellite data can become an alternative and supplement to official reporting, as well as be used for control and surveillance activities and bringing to legal responsibility. Topical issues are publishing official environmental information in a form of open data, the development of mechanisms for public non-financial reporting of large companies, the study of the links between information transparency and reliability of the information, as well as the peculiarities of the formation of official reporting about environmental impact."
keywords: "*Keywords:* official ecological information, right for information, openness of information, availability of information, reliability of inrofmation, remote sensing of environment, air pollution, sulphur dioxide, space monitoring."
output:
  odt_document:
    reference_odt: reference-styles.odt
    template: template.opendocument
bibliography: references.bib
csl: spbu-gerald-law-for-references.csl
notes-after-punctuation: false
reference-section-title: "References"
nocite: |
  @*
---